% A LaTeX template for MSc Thesis submissions to 
% Politecnico di Milano (PoliMi) - School of Industrial and Information Engineering
%
% S. Bonetti, A. Gruttadauria, G. Mescolini, A. Zingaro
% e-mail: template-tesi-ingind@polimi.it
%
% Last Revision: October 2021
%
% Copyright 2021 Politecnico di Milano, Italy. NC-BY

\documentclass{Configuration_Files/PoliMi3i_thesis}

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------

% CONFIGURATIONS
\usepackage{parskip} % For paragraph layout
\usepackage{setspace} % For using single or double spacing
\usepackage{emptypage} % To insert empty pages
\usepackage{multicol} % To write in multiple columns (executive summary)
\setlength\columnsep{15pt} % Column separation in executive summary
\setlength\parindent{10pt} % Indentation
\raggedbottom  

% PACKAGES FOR TITLES
\usepackage{titlesec}
% \titlespacing{\section}{left spacing}{before spacing}{after spacing}
\titlespacing{\section}{0pt}{3.3ex}{2ex}
\titlespacing{\subsection}{0pt}{3.3ex}{1.65ex}
\titlespacing{\subsubsection}{0pt}{3.3ex}{1ex}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[english]{babel} % The document is in English  
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding
\usepackage[11pt]{moresize} % Big fonts

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\usepackage{svg}
\usepackage{transparent} % Enables transparent images
\usepackage{eso-pic} % For the background picture on the title page
%\usepackage{subfig} % Numbered and caption subfigures using \subfloat.
\usepackage[export]{adjustbox}
\usepackage{tikz} % A package for high-quality hand-made figures.
\usetikzlibrary{}
\graphicspath{{./Images/}} % Directory of the images
\usepackage{caption} % Coloured captions
\usepackage{subcaption}
\usepackage{xcolor} % Coloured captions
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{float}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{cancel}
\usepackage[overload]{empheq} % For braced-style systems of equations.
\usepackage{fix-cm} % To override original LaTeX restrictions on sizes

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % Tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
\bibliographystyle{abbrvnat} % You may use a different style adapted to your field

% OTHER PACKAGES
\usepackage{pdfpages} % To include a pdf file
\usepackage{afterpage}
\usepackage{lipsum} % DUMMY PACKAGE
\usepackage{fancyhdr} % For the headers
\usepackage{fancyvrb}
\usepackage[acronym]{glossaries}
\usepackage{enumitem} 
\fancyhf{}

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{placeins}


% Input of configuration file. Do not change config.tex file unless you really know what you are doing. 
\input{Configuration_Files/config}

%----------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%----------------------------------------------------------------------------

% EXAMPLES OF NEW COMMANDS
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation

%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------

%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------

\input{Thesis_Acronyms}

%----------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%----------------------------------------------------------------------------

\begin{document}

\fancypagestyle{plain}{%
\fancyhf{} % Clear all header and footer fields
\fancyhead[RO,RE]{\thepage} %RO=right odd, RE=right even
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%----------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\puttitle{
	title={VR-Enabled Social Robotics: Development and Evaluation of an Immersive Teleoperation Platform for Human-Robot Interaction}, % Title of the thesis
	name=Sebastian Enrique Perea Lopez, % Author Name and Surname
	course=Computer Science and Engineering \\Ingegneria Informatica, % Study Programme (in Italian)
	ID  =  10986638,  % Student ID number (numero di matricola)
	advisor= Prof. Andrea Bonarini, % Supervisor name
	coadvisor={Federico Espositi}, % Co-Supervisor name, remove this line if there is none
	academicyear={2024-25},  % Academic Year
} % These info will be put into your Title page 

%----------------------------------------------------------------------------
%	PREAMBLE PAGES: ABSTRACT (inglese e italiano), EXECUTIVE SUMMARY
%----------------------------------------------------------------------------

\pagebreak
\pagestyle{empty}
\hspace{0pt}
\vfill
\textit{Dedicado a mi familia.}
\vfill
\hspace{0pt}
\pagebreak

\startpreamble
\setcounter{page}{1} % Set page counter to 1

% ABSTRACT IN ENGLISH
\chapter*{Abstract} 
This thesis presents the development of Tino V2, a social robotics platform designed for immersive Virtual Reality (VR) teleoperation while preserving the expressive movement capabilities required for natural human-robot interaction. Building upon the limitations of a legacy Raspberry Pi-based system, Tino V2 introduces a modern distributed architecture that integrates advanced sensing and real-time AI processing.

The system combines Ultra-Wideband (UWB) positioning with RTABMap visual SLAM to achieve a 4.7× improvement in localization accuracy, while real-time human perception is enabled through a YOLOv11-based 3D pose estimation pipeline operating at sub-24 ms latency. A ROS2–Unity communication framework supports natural VR-based robot control with 98\% command reliability, while distributed execution on NVIDIA Orin Nano enables concurrent SLAM, perception, and sensor fusion processes with 93\% uptime.

Evaluation with twelve participants in collaborative maze scenarios demonstrated that immersive VR control allows effective spatial coordination, non-verbal communication, and problem-solving, confirming that teleoperation can sustain the social qualities essential for meaningful human-robot interaction.

This work contributes novel methods for hybrid localization, VR-optimized movement primitives, and a comprehensive framework for evaluating VR-mediated social interaction. Results establish immersive teleoperation as a viable paradigm for social robotics research, laying the groundwork for future studies in telepresence, empathy transfer, and distributed collaboration.\\
\\
\textbf{Keywords:} social robotics, virtual reality teleoperation, human-robot interaction, SLAM, pose estimation, sensor fusion, ROS2% Keywords

% ABSTRACT IN ITALIAN
\chapter*{Abstract in lingua italiana}
Questa tesi presenta lo sviluppo di Tino V2, una piattaforma di robotica sociale progettata per la teleoperazione immersiva in Realtà Virtuale (VR) mantenendo le capacità di movimento espressive necessarie per un'interazione naturale uomo-robot. Basandosi sulle limitazioni di un sistema legacy basato su Raspberry Pi, Tino V2 introduce un'architettura distribuita moderna che integra sensori avanzati e elaborazione AI in tempo reale.

Il sistema combina il posizionamento Ultra-Wideband (UWB) con RTABMap visual SLAM per ottenere un miglioramento di 4,7× nella precisione di localizzazione, mentre la percezione umana in tempo reale è abilitata attraverso un pipeline di stima della posa 3D basata su YOLOv11 che opera con una latenza inferiore a 24 ms. Un framework di comunicazione ROS2–Unity supporta il controllo naturale del robot basato su VR con una affidabilità del comando del 98\%, mentre l'esecuzione distribuita su NVIDIA Orin Nano consente processi simultanei di SLAM, percezione e fusione dei sensori con un uptime del 93\%.

La valutazione con dodici partecipanti in scenari collaborativi di labirinti ha dimostrato che il controllo immersivo in VR consente una coordinazione spaziale efficace, comunicazione non verbale e problem-solving, confermando che la teleoperazione può sostenere le qualità sociali essenziali per un'interazione significativa uomo-robot.

Questo lavoro contribuisce con metodi innovativi per la localizzazione ibrida, primitive di movimento ottimizzate per la VR e un framework completo per la valutazione dell'interazione sociale mediata dalla VR. I risultati stabiliscono la teleoperazione immersiva come un paradigma praticabile per la ricerca sulla robotica sociale, ponendo le basi per futuri studi sulla telepresenza, il trasferimento di empatia e la collaborazione distribuita.\\
\\
\textbf{Parole chiave:} robotica sociale, teleoperazione in realtà virtuale, interazione uomo-robot, SLAM, stima della posa, fusione dei sensori, ROS2% Keywords (italian)

%----------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES/SYMBOLS
%----------------------------------------------------------------------------

% TABLE OF CONTENTS
\thispagestyle{empty}
\tableofcontents % Table of contents 
\thispagestyle{empty}
\cleardoublepage

%-------------------------------------------------------------------------
%	THESIS MAIN TEXT
%-------------------------------------------------------------------------
% In the main text of your thesis you can write the chapters in two different ways:
%
%(1) As presented in this template you can write:
%    \chapter{Title of the chapter}
%    *body of the chapter*
%
%(2) You can write your chapter in a separated .tex file and then include it in the main file with the following command:
%    \chapter{Title of the chapter}
%    \input{chapter_file.tex}
%
% Especially for long thesis, we recommend you the second option.


\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering


% Chapter 1: Introduction
\chapter{Introduction}
\input{Chapters/introduction}

% Chapter 2: Background
\chapter{Background}
\input{Chapters/background}

% Chapter 3: Conceptual Work
\chapter{Conceptual Work}
\input{Chapters/conceptual_work}

% Chapter 4: Implementation
\chapter{Implementation}
\input{Chapters/implementation}

\input{Chapters/Hardware_implementation}

\input{Chapters/Software_implementation}

\input{Chapters/Perception_implementation}

\input{Chapters/Control_implementation}

\input{Chapters/Integration_implementation.tex}


% Chapter 5: Evaluation
\chapter{System Evaluation and Validation}
\input{Chapters/evaluation}

% Chapter 6: Conclusions
\chapter{Conclusions}
\input{Chapters/conclusions}

% \input{Chapters/temporalRD}

% --------------------------------------------------------------------------
% NUMBERED CHAPTERS % Regular chapters following
% --------------------------------------------------------------------------

%-------------------------------------------------------------------------
%	BIBLIOGRAPHY
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\bibliography{Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

%-------------------------------------------------------------------------
%	APPENDICES
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
% \appendix
% \chapter{Appendix A}
% If you need to include an appendix to support the research in your thesis, you can place it at the end of the manuscript.
% An appendix contains supplementary material (figures, tables, data, codes, mathematical proofs, surveys, \dots)
% which supplement the main results contained in the previous chapters.

% \chapter{Appendix B}
% It may be necessary to include another appendix to better organize the presentation of supplementary material.
\printglossary[type=\acronymtype]

% LIST OF FIGURES
\listoffigures

% LIST OF TABLES
\listoftables

% % LIST OF SYMBOLS
% % Write out the List of Symbols in this page
% \chapter*{List of Symbols} % You have to include a chapter for your list of symbols (
% \begin{table}[H]
%     \centering
%     \begin{tabular}{lll}
%         \textbf{Variable} & \textbf{Description} & \textbf{SI unit} \\\hline\\[-9px]
%         $\bm{u}$ & solid displacement & m \\[2px]
%         $\bm{u}_f$ & fluid displacement & m \\[2px]
%     \end{tabular}
% \end{table}

% ACKNOWLEDGEMENTS
\chapter*{Acknowledgements}

This thesis represents not just my individual effort, but the culmination of support, guidance, and collaboration from many wonderful people who have made this journey possible and meaningful.

First and foremost, I extend my deepest gratitude to Professor Andrea Bonarini, whose guidance, wisdom, and patience have been instrumental throughout this entire process. His passion for robotics and human-robot interaction has been truly inspiring, and his mentorship has shaped not only this work but also my perspective as a researcher. I am equally grateful to Federico Espositi, whose technical expertise, insightful feedback, and continuous support have been invaluable in navigating the complex challenges of this project.

To the entire AIRLab community—thank you for creating such a stimulating and collaborative environment where learning and innovation flourish. The knowledge I gained, the discussions we shared, and the friendships formed within these walls have enriched both my research and my personal growth in ways I could never have imagined.

A special acknowledgment goes to Alessandro, whose brilliant work on the VR development made it possible to bridge two entirely different worlds—the physical and the virtual. Without his dedication and technical prowess, the core vision of this thesis could never have been realized. Our collaboration demonstrated the true power of interdisciplinary teamwork.

To my parents, Alexandra and Jorge—words cannot express how grateful I am for your unwavering love and support. Despite the 7-hour time difference that separated us, you were always there for me during the most challenging moments of this journey. Your sacrifices, encouragement, and belief in me have been the foundation upon which all my achievements rest. Without your tireless efforts and dedication throughout my life, I would never have reached this milestone.

To my incredible friends Nicole, Daniel, and Camilo—thank you for your patience, humor, and enthusiasm during this adventure. You not only helped me test Tino through countless experiments but also witnessed me doing some truly crazy things during development (and somehow still remained my friends!). Your friendship, laughter, and willingness to be part of this research made even the most stressful moments enjoyable and memorable.

Finally, I want to acknowledge everyone who contributed to this work in both big and small ways—from technical discussions to moral support, from debugging sessions to celebration moments. This thesis is as much yours as it is mine.

Thank you all for being part of this incredible journey.

\cleardoublepage

\end{document}
